---
title: "20160915_Thompson_HW-01"
author: "Nick Thompson"
date: "September 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Method of MLE

Find the Maximum Likelihood Estimate for $\pi$.

* The probability model of the binomial distribution is:  
$$f(y|n,\pi)={{n}\choose{y}}\pi^y(1-\pi)^n-y$$

* Step 1:  Write down the likelihood function (I did this step for you):
$$\mathcal{L}(\pi|y,n)=p(y|n,\pi) = {{n}\choose{y}}\pi^y(1-\pi)^n-y$$

* Step 2:  Take the log and reduce
$$\mathcal{L}(\pi|y,n) = \sum_{i=1}^n log \pi^y * \sum_{i=1}^n log(1-\pi)^{n-y}$$
$$\mathcal{L}(\pi|y,n) = \sum_{i=1}^n y\ log \pi * \sum_{i=1}^n (n-y)\ log(1-\pi)$$

* Step 3:  Maximize and find the most likely value for $\pi$ (Start with the log-likelihood and take the derivative with respect to the parameter).

$$\frac{d\pi}{dy} \mathcal{L} = n\bar{y}\log\pi + (n-\bar{y})n\log(1-\pi) \rightarrow$$


# R

The maximizing function we used in class used the "seq" function to generate a sequence from .01 to .99 in increments of .01.

```{r p.seq}
#this command
p.seq <- seq(0.01, 0.99, 0.01)
```

* How would you change the code to create a sequence from .01 to .99 in increments of .1?

```{r p.seq2, echo=TRUE}
p.seq2 <- seq(0.01, 0.99, 0.1)
```

* How would you get R to print the first ten items in the sequence you just created? (look at subsetting)

```{r p.seq2_sub, echo=TRUE}
p.seq2_sub <- p.seq2[1:10]
```
